Inferring generation-interval distributions from contact tracing data: \\ \emph{in prep, PRSB}

### 2018-10-09 doc/network.tex
trying to make things a little clearer

We begin by introducing an analytical framework.
Let $K_a(\tau)$ be an individual level kernel, i.e., the rate at which secondary infections are caused by an individual $a$ \citep{svensson2015influence}. \swp{Need to say something more about Sven; come back later}
Then, the unconditional, population-level kernel is given by integrating over individual variations:
\begin{equation}
K(\tau) = \int K_a (\tau) dA.
\end{equation}
The basic reproductive number -- average number of secondary cases caused by an average primary case in a fully susceptible population -- is defined as: 
\begin{equation}
\RR_0 = \int K(t).
\end{equation}
Then, the intrinsic generation-interval is simply the population-level kernel normalized by the basic reproductive number:
\begin{equation}
g(\tau) = \frac{K(t)}{\RR_0}.
\end{equation}
Then, the renewal equation model can be used to model incidence over time:
\begin{equation}
i(\tau) = S(\tau) \int K(s) i(\tau-s) ds = \RR \int g(s) i(\tau-s) ds
\end{equation}
where $i(\tau)$ is incidence at time $\tau$ and $S(\tau)$ is the proportion of the population susceptible.
Finally, assuming exponentially growing incidence ($i(t) = i(0) \exp(r t)$), we obtain the Euler-Lotka equation [CITE]:
\begin{equation}
\frac{1}{\RR} = \int g(\tau) \exp(-r \tau) d\tau.
\end{equation}

\subsection*{Right-censored intervals}

@@

\subsection*{Conditional kernel}

Given limited contact structure, a susceptible individual can be contacted multiple times;
infectious contacts give rise to infection only when the susceptible individual is contacted for the first time.
Then, the conditional kernel (conditional on the infection occuring) of an individual $a$ is given by
\begin{equation}
\hat{k}_a(\tau) = k_a(\tau) \exp \left(- \delta_a \int_0^\tau k_a(s) ds\right),
\end{equation}
where $\delta_a$ is the dilution.
The individual kernel, $k_a(s)$, measures the intrinsic infectiousness of an individual $a$,
whereas $\delta_a k_a(s)$ measures the infectiousness of an individual $a$ for a particular susceptible individual.
Then, the population-level conditional kernel is:
\begin{equation}
\hat{k}(\tau) = \int \hat{k}_a(\tau) dA,
\end{equation}
and the conditional generation-interval distribution, defined to be the distribution of times at which secondary infections occur from a single primary infector, is:
\begin{equation}
\hat{g}(\tau) = \frac{\hat{k}(\tau)}{\int \hat{k} \tau}.
\label{eq:conditional}
\end{equation}
\swp{Need to say something about trapman. Maybe say this is a generalization of Trapman's work...? Is that too much? Do we need to give an example of what conditional kernels look like?}

The difference between conditional and intrinsic generation-interval distribution can be demonstrated by simulating stochastic infection processes on a small tree network (i.e, a single infected individual is connected to multiple susceptible individuals who are not connected with each other) with artificially high $\RR$.
Simulations confirm that in this case the distribution of contact times matches the intrinsic generation-interval distribution, while the distribution of generation times matches with the conditional generation-interval distribution (Figure~\ref{fig:local}) -- generation times are shorter on average because contacts with already-infected susceptibles do not lead to new infections.

The conditional generation interval \eqref{conditional} only explains some of the reduction in generation times that occurs on most networks, however.
Generation intervals are also shortened by indirect connections: if a susceptible individual is infected through another route while the focal individual is infected, this will shorten the generation interval \jd{This s.~is too repetitive.}
Simulations on a small homogeneous network confirm this additional effect (\fref{local}, right panel). 
In general, estimating the effects of most networks on observed generation intervals is expected to be a hard problem.

@@

>>>>

\subsection{Intrinsic kernel}

We begin by introducing an analytical framework.
Let $K_a(\tau)$ be an individual-level intrinsic kernel, i.e., the rate at which secondary infections are expected to be caused by an individual $a$ \citep{svensson2015influence}; \swp{Need to say something more about Sven; come back later}
an individual-level intrinsic kernel describes an intrinsic infectiousness of an infected individual.
The population-level intrinsic kernel is given by integrating over individual variations:
\begin{equation}
K(\tau) = \int K_a (\tau) dA.
\end{equation}
The basic reproductive number -- average number of secondary cases caused by an average primary case in a fully susceptible population -- is defined as: 
\begin{equation}
\RR_0 = \int K(t).
\end{equation}
The intrinsic generation-interval is simply the population-level kernel normalized by the basic reproductive number:
\begin{equation}
g(\tau) = \frac{K(t)}{\RR_0}.
\end{equation}
Then, the renewal equation model can be used to model incidence over time:
\begin{equation}
i(\tau) = S(\tau) \int K(s) i(\tau-s) ds = \RR \int g(s) i(\tau-s) ds
\end{equation}
where $i(\tau)$ is incidence at time $\tau$ and $S(\tau)$ is the proportion of the population susceptible.
Finally, assuming exponentially growing incidence ($i(t) = i(0) \exp(r t)$), we obtain the Euler-Lotka equation [CITE]:
\begin{equation}
\frac{1}{\RR} = \int g(\tau) \exp(-r \tau) d\tau.
\end{equation}

\subsection{Right-censored intervals}

@@

\subsection{Egocentric kernel}

Within a limited contact structure, a susceptible individual can be contacted multiple times;
infectious contacts give rise to infection only when the susceptible individual is contacted for the first time.
In particular, we can look at it from an egocentric point of view -- i.e., from a perspective of a single infector.
Then, we can define the egocentric kernel, i.e., the rate at which secondary infections are realized by a single primary case $a$:
\begin{equation}
\hat{K}_a(\tau) = K_a(\tau) \exp \left(- \delta_a \int_0^\tau K_a(s) ds\right),
\end{equation}
where $K_a(\tau)$ is the individual-level intrinsic kernel and $e^{- \delta_a \int_0^\tau K_a(s) ds}$ is the probability that a susceptible acquaintance has not yet been contacted by individual $a$.
The dilution term, $\delta_a$, models how contacts are distributed among susceptible acquaintances.
For example, when infectious contacts are distributed equally in a homogeneous population, we have $\delta_a = 1/(N-1)$, where $N$ is the population size.

The population-level egocentric kernel is given by integrating over individual variations:
\begin{equation}
\hat{K}(\tau) = \int \hat{K}_a(\tau) dA,
\end{equation}
and the population-level egocentric generation-interval distribution is:
\begin{equation}
\hat{g}(\tau) = \frac{\hat{K}(\tau)}{\int \hat{K} \tau}.
\label{eq:conditional}
\end{equation}
The population-level egocentric generation-interval distribution describes the distribution of times at which secondary infections are realized by an average primary case.
For convenience, egocentric distribution refers to the population-level distribution, unless noted otherwise.

For example, consider a susceptible-exposed-infected-recovered (SEIR) model.
This model assumes that latent and infectious periods are exponentially distributed and can describe wide range of diseases.
It has been shown that the intrinsic generation-interval distribution that corresponds to this model is given by [CITE]:
\begin{equation}
g(\tau) = \frac{\sigma \gamma}{\sigma - \gamma} \left(e^{-\gamma t} - e^{-\sigma t}\right),
\end{equation}
where $1/\sigma$ and $1/\gamma$ are mean latent and infectious periods, respectively.
Assuming that per-pair contact rate is $\lambda$ for any pair, we obtain the following egocentric generation-interval distribution is given by:
\begin{equation}
\hat{g}(\tau) = \frac{\sigma (\gamma + \lambda)}{\sigma - (\gamma + \lambda)} \left(e^{-(\gamma + \lambda)t} - e^{-\sigma t}\right)
\end{equation}
In this scenario, per-pair contact rate is effectively decreasing mean infectious period: an infected individual is can no longer infect anyone (hence no longer infectious) when all its susceptible acquaintances have ben infected.

The difference between egocentric and intrinsic generation-interval distribution can be demonstrated by simulating stochastic infection processes on a small tree network (i.e, a single infected individual is connected to multiple susceptible individuals who are not connected with each other) with artificially high per-pair contact rate.
Simulations confirm that in this case the distribution of contact times matches the intrinsic generation-interval distribution, while the distribution of realized generation times matches with the egocentric generation-interval distribution (\fref{fig:local}) -- generation times are shorter on average because contacts with already-infected susceptibles do not lead to new infections.

@@

The egocentric generation interval \eqref{conditional} only explains some of the reduction in generation times that occurs on most networks, however.
Generation intervals are also shortened by indirect connections: a susceptible individual can be infected through another route before the focal individual makes infectious contacts.
Simulations on a small homogeneous network confirm this additional effect (\fref{local}, right panel). 
In general, estimating the effects of most networks on observed generation intervals is expected to be a hard problem.

\subsection{Linking growth rate and reproductive number}
